---
title: 'CPO'
output:
  mikedown::ucsb:
    toc: TRUE
    number_sections: FALSE
---
# Explority data analysis
# Angus Watters
# 06/09/2021

```{r, warning=FALSE, message=FALSE}
# libraries
library(tidyverse)
library(tidymodels) # modeling
library(jtools) # Linear regression exploration
library(vip) # Variable importance 
library(sensemakr) # sensitivty analysis 
library(psych) # correlation functions
```

```{r, warning=FALSE, message=FALSE}
df <- readRDS("C:/Users/angus/OneDrive/Desktop/lynker/CPO/data/model_data/model_data_non_spatial.rds")
```

```{r, warning=FALSE, message=FALSE}
df <- df %>% 
  mutate(month = lubridate::month(date), year = lubridate::year(date)) %>% 
  mutate(season = case_when(
    month %in% c(4, 5, 6, 7, 8, 9) ~ "growing",
    month %in% c(1, 2, 3, 10, 11, 12) ~"not_growing"
  )
  ) %>% 
  group_by(year, season, district) %>% # calculate averages during growing and not growing seasons
  summarize(
    short_mean = mean(short),
    prcp_mean = mean(prcp),
    soilm_mean = mean(soilm),
    pdsi_mean = mean(pdsi_terra),
    tavg_mean = mean(tavg_c),
    swe_max = max(swe), 
    spi1_mean = mean(spi1),
    spi12_mean = mean(spi12),
    water_deficit_mean = mean(water_deficit),
    aet_mean = mean(aet)
  ) 

df$season <- as.factor(df$season)
# seperate dataframes into growing and not growing, widen dataframe and make average columns for growing and not growing
grow <- df %>% 
  filter(season == "growing") %>% 
  pivot_wider(names_from = "season", values_from = c(short_mean:aet_mean))
not_grow <- df %>% 
  filter(season == "not_growing", year > 1979) %>% 
  pivot_wider(names_from = "season", values_from = c(short_mean:aet_mean))

# join wider growing and not growing season dataframes 
season_shorts <- inner_join(grow, not_grow, by = c("district", "year"))

```

```{r}
season_shorts %>%
  mutate(district = as.character(district)) %>% 
  ggplot(aes(fct_reorder(district, short_mean_growing),
             short_mean_growing,
             fill = district
  )) +
  geom_boxplot(outlier.alpha = 0.5) +
  coord_flip() +
  labs(
    fill = NULL, x = "District",
    y = "Growing season mean shortage"
  )
```

```{r}
df %>%
  mutate(district = as.character(district)) %>% 
  ggplot(aes(fct_reorder(district, short_mean),
             short_mean,
             fill = season
  )) +
  geom_boxplot(outlier.alpha = 0.5) +
  coord_flip() +
  labs(
    fill = NULL, x = "District",
    y = "Seasonal shortage mean"
  )
```

# Correlation

## Correlation matrix
```{r}
# isolate mean shortage during growing season
growing_season <- dplyr::select(season_shorts, -short_mean_not_growing) 

DT::datatable(print(corr.test(growing_season,
          use    = "pairwise",
          method = "pearson",
          adjust = "none"), short = FALSE))
```


##  Variable correlation
```{r}
pairs(data = growing_season, short_mean_growing~spi1_mean_growing + spi1_mean_not_growing + swe_max_growing + swe_max_not_growing + tavg_mean_growing+  tavg_mean_not_growing)
```

##  Correlation histograms

### Growing season correlations
```{r}
grow_corr <- growing_season[,3:12]
PerformanceAnalytics::chart.Correlation(grow_corr,
                   method="pearson",
                   histogram=TRUE,
                   pch=16)
```

### Non growing season correlations
```{r}
nongrow_corr <- growing_season[,13:21]
PerformanceAnalytics::chart.Correlation(nongrow_corr,
                   method="pearson",
                   histogram=TRUE,
                   pch=16)
```

# *Linear regression model*
```{r}
# linear model specifications
lm_fit <- lm(short_mean_growing~., data = growing_season)
```


## Extract Variable importance scores
```{r}
vip::vip(lm_fit, 
          mapping = aes_string(fill = "Sign"),
    aesthetics = list(color = "black"), num_features = 20) 
```

## Coefficient summary
```{r}
jtools::summ(lm_fit, scale= TRUE, center = TRUE)
```

## **Effect plots:** regression estimates
```{r, warning=FALSE, message=FALSE}
p1 <- jtools::effect_plot(lm_fit, pred = spi1_mean_growing,  interval = TRUE, plot.points = TRUE)
p2 <- jtools::effect_plot(lm_fit, pred = spi1_mean_not_growing,  interval = TRUE, plot.points = TRUE)
p3 <- jtools::effect_plot(lm_fit, pred = tavg_mean_growing,  interval = TRUE, plot.points = TRUE)
p4 <- jtools::effect_plot(lm_fit, pred = tavg_mean_not_growing,  interval = TRUE, plot.points = TRUE)
p5 <- jtools::effect_plot(lm_fit, pred = spi12_mean_growing,  interval = TRUE, plot.points = TRUE)
p6 <- jtools::effect_plot(lm_fit, pred = spi12_mean_not_growing,  interval = TRUE, plot.points = TRUE)
p7 <- jtools::effect_plot(lm_fit, pred = swe_max_growing,  interval = TRUE, plot.points = TRUE)
p8 <- jtools::effect_plot(lm_fit, pred = swe_max_not_growing,  interval = TRUE, plot.points = TRUE)
p9 <- jtools::effect_plot(lm_fit, pred = prcp_mean_growing,  interval = TRUE, plot.points = TRUE)
p10 <- jtools::effect_plot(lm_fit, pred = prcp_mean_not_growing,  interval = TRUE, plot.points = TRUE)
p11 <- jtools::effect_plot(lm_fit, pred = pdsi_mean_growing,  interval = TRUE, plot.points = TRUE)
p12 <- jtools::effect_plot(lm_fit, pred = pdsi_mean_not_growing,  interval = TRUE, plot.points = TRUE)
cowplot::plot_grid(p1, p2, p3, p4, p5, p6,  label_size = 10)
```

```{r, warning=FALSE, message=FALSE}
cowplot::plot_grid(p7, p8, p9, p10, p11, p12, label_size = 10)
```

# Regression coefficient plots
```{r, warning=FALSE, message=FALSE}
plot_summs(lm_fit, plot.distributions = TRUE, scale = TRUE, colors = 'Rainbow')
# plot_coefs(lm_fit, scale = TRUE)
```

# Coefficient uncertainty as normal distributions
```{r, warning=FALSE, message=FALSE}
plot_summs(lm_fit, scale = TRUE, plot.distributions = TRUE, inner_ci_level = .9)
```


# Sensitivity analysis
```{r}
short_sensitivity <- sensemakr::sensemakr(model = lm_fit, 
                                treatment = "spi1_mean_growing",
                                benchmark_covariates = "district",
                                kd = 1:3,
                                ky = 1:3, 
                                q = 1,
                                alpha = 0.05, 
                                reduce = TRUE)
plot(short_sensitivity)
```


















